### 2014-11-07 myFigureWnorm.tex

        %%%%%%%%%%% Example 1 %%%%%%%%%%%%%%%
        \begin{minipage}{1.0\linewidth}
        \end{minipage}
        % QUERY image
        \begin{minipage}{0.34\linewidth}
            \centering
            \vspace{0mm}
            \includegraphics[trim = 45mm 40mm 45mm 30mm, clip=true,height=36mm]{imgs/Wnorm/exImproved02/query.jpg}
        \end{minipage}
        % Retrieved images
        \begin{minipage}{0.75\linewidth}
            % FV e-SVM
            \begin{minipage}{\linewidth} 
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improvedPval01.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improvedPval02.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improvedPval03.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improvedPval04.jpg}}
            \end{minipage}
            \\
            \begin{minipage}{\linewidth}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improved01.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improved02.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improved03.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved02/improved04.jpg}}
            \end{minipage} 
        \end{minipage}
        \vspace{3mm}
        \\
        %%%%%%%%%%% Example 2 %%%%%%%%%%%%%%%
        \begin{minipage}{1.0\linewidth}
        \end{minipage}
        % QUERY image
        \begin{minipage}{0.34\linewidth}
            \centering
            \vspace{0mm}
            \includegraphics[trim = 45mm 40mm 45mm 30mm, clip=true,height=36mm]{imgs/Wnorm/exImproved05/query.jpg}
        \end{minipage}
        % Retrieved images
        \begin{minipage}{0.75\linewidth}
            % FV e-SVM
            \begin{minipage}{\linewidth} 
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improvedPval01.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improvedPval02.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improvedPval03.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improvedPval04.jpg}}
            \end{minipage}
            \\
            \begin{minipage}{\linewidth}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improved01.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improved02.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improved03.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved05/improved04.jpg}}
            \end{minipage} 
        \end{minipage}
        \vspace{3mm}
        \\
        %%%%%%%%%%% Example 3 %%%%%%%%%%%%%%%
        \begin{minipage}{0.34\linewidth}
            \centering
            \vspace{0mm}
            \includegraphics[trim = 45mm 40mm 45mm 30mm, clip=true, height=36mm]{imgs/Wnorm/exImproved04/query.jpg}
        \end{minipage}
        % Retrieved images
        \begin{minipage}{0.75\linewidth}
            % FV e-SVM
            \begin{minipage}{\linewidth} 
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improvedPval01.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improvedPval02.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improvedPval03.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improvedPval04.jpg}}
            \end{minipage}
            \\
            \begin{minipage}{\linewidth}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improved01.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improved02.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improved03.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exImproved04/improved04.jpg}}
            \end{minipage} 
        \end{minipage}
        \vspace{3mm}
        \\
        %%%%%%%%%%% MIX FAILURE CASES: good examples are Mix09 14 17 16 20
        %%%%%%%%%%% Example 4 %%%%%%%%%%%%%%%
        \begin{minipage}{0.34\linewidth}
            \centering
            \vspace{0mm}
            \includegraphics[trim = 45mm 40mm 45mm 30mm, clip=true, height=36mm]{imgs/Wnorm/exMix15/query.jpg}
        \end{minipage}
        % Retrieved images
        \begin{minipage}{0.75\linewidth}
            % FV e-SVM
            \begin{minipage}{\linewidth} 
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mixPval01.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mixPval02.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mixPval03.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mixPval04.jpg}}
            \end{minipage}
            \\
            \begin{minipage}{\linewidth}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mix01.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mix02.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mix03.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix15/mix04.jpg}}
            \end{minipage} 
        \end{minipage}
        \vspace{3mm}
        \\
        %%%%%%%%%%% Example 5 %%%%%%%%%%%%%%%
        \begin{minipage}{0.34\linewidth}
            \centering
            \vspace{0mm}
            \includegraphics[trim = 45mm 40mm 45mm 30mm, clip=true, height=36mm]{imgs/Wnorm/exMix19/query.jpg}
        \end{minipage}
        % Retrieved images
        \begin{minipage}{0.75\linewidth}
            % FV e-SVM
            \begin{minipage}{\linewidth} 
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mixPval01.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mixPval02.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mixPval03.jpg}}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mixPval04.jpg}}
            \end{minipage}
            \\
            \begin{minipage}{\linewidth}
                \colorbox{myGreen}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mix01.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mix02.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mix03.jpg}}
                \colorbox{myRed}{\includegraphics[trim = 35mm 30mm 35mm 30mm, clip=true, height=16mm]{imgs/Wnorm/exMix19/mix04.jpg}}
            \end{minipage} 
        \end{minipage}
        %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%