### 2014-11-20 supplementary/supp.tex
Supplementary material

\documentclass[10pt,onecolumn,A4]{article}

\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% usepackages by Petr
\usepackage{epstopdf}
\usepackage{enumerate}
\usepackage{subfigure}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[table]{xcolor}
\usepackage{hyphenat}
\setlength\textwidth{175mm}
\setlength\textheight{280mm}
\setlength\oddsidemargin{0mm}
\setlength\evensidemargin{0mm}
\setlength\topmargin{-25mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
\setlength\leftmargin{0mm}

\definecolor{myGrey}{HTML}{33FF00}
\definecolor{myRed}{HTML}{FF3030}
\definecolor{myGreen}{HTML}{00FF00}
\definecolor{myGrey}{HTML}{AAAAAA}
\definecolor{myWhite}{HTML}{FFFFFF}
\definecolor{myCopper1}{HTML}{FFC77F}
\definecolor{myCopper2}{HTML}{FFAB6D}
\definecolor{myCopper3}{HTML}{E9915D}
\definecolor{myCopper4}{HTML}{BB754B}
\definecolor{myCopper5}{HTML}{935C3A}

\title{Supplementary material, paperID:954}
\begin{document}
	\maketitle

\begin{figure}
	\center
	\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=0.5\linewidth]{mapGoogle.png}
	\caption{ \textbf{A map corresponding to the area covered by database.}
		The Pittsburgh 25k dataset covers roughly an area of $1.3 \times 1.2 \; {\rm km}^2$.
	}
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1383/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1383/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/svm01.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1383/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%\caption{Example sup1383} 
	\caption{ \textbf{Response of Fisher vector accross the databse images.}
		\emph{Top row:} The heatmaps show a distribution of the Fisher vector (FV) similarity score with a given query image. A location of the query image is shown by green circle, top 5 strongest peaks are denoted by a cross marker along with its rank. \emph{Bottom row:} The query image is shown in a gray box and top five database images corresponding to the top 5 strongest peaks in the map. The correct match is denoted by green bounding box.
		Each FV score votes into a max-pooling accumulation array that corresponds to a GPS location of the database images.
		\emph{Left} column shows results for raw FV baseline without any learning while \emph{right} column displays results for FV obtained by proposed method (e-SVM FV). Notice that e-SVM FV descriptor suppress the peaks that are far from the query image which results into a change of ranking.
		}
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1666/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1666/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1666/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%\caption{Example sup1666} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1693/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1693/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1693/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup1693} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1812/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1812/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1812/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup1812} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1957/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1957/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1957/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup1957} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1962/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup1962/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/svm01.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup1962/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup1962} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2358/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2358/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/svm03.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2358/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup2358} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2569/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2569/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/svm03.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2569/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup2569} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2602/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2602/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/svm01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/svm03.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2602/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup2602} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2959/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2959/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/svm01.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2959/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup2959} 
\end{figure}

\begin{figure}
	\begin{minipage}{0.45\linewidth}
		\center
		(raw FV128 baseline) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2961/heatRaw.jpg}
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\center
		(e-SVM FV128) \\
		\includegraphics[trim = 55mm 40mm 55mm 25mm, clip=true,width=\linewidth]{sup2961/heatSvm.jpg}
	\end{minipage} 
	\\
	\textcolor{myWhite}{xxx}\\
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/raw01.jpg}}
		\colorbox{myCopper2}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/raw02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/raw03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/raw04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/raw05.jpg}} \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage} 
	\begin{minipage}{0.45\linewidth}
		\colorbox{myGrey}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/query.jpg}}
		\colorbox{myCopper1}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/svm01.jpg}}
		\colorbox{myGreen}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/svm02.jpg}}	\\
		\textcolor{myWhite}{xxxxxx}Query \hspace{0.25\linewidth}1. \hspace{0.25\linewidth}2. \\
		\colorbox{myCopper3}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/svm03.jpg}}
		\colorbox{myCopper4}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/svm04.jpg}}
		\colorbox{myCopper5}{\includegraphics[trim = 55mm 40mm 55mm 30mm, clip=true,width=0.30\linewidth]{sup2961/svm05.jpg}}  \\
		\textcolor{myWhite}{xxxxxx}3. \hspace{0.25\linewidth}4. \hspace{0.25\linewidth}5. \\
	\end{minipage}
	%%\caption{Example sup2961} 
\end{figure}

\end{document}